### 2013-10-15 Writing/paper.tex
Update after a long absence… still not quite done

%Use this template for comments in the margin: \com{Note!}

%\date{December 2010}                           % Activate to display a given date or no date

%
%  From Riggle:
%
%----------------------------------------------------------------------------%
%                    Basic page layout parameters                            %
%----------------------------------------------------------------------------%

% \documentclass[11pt,leqno,tbtags]{article}
\documentclass[12pt,leqno,tbtags,twoside]{article}

    % leqno = equations at the left
    % tbtags (+leqno) = equation numbers at the top (rather than centered)

\usepackage[letterpaper,hmargin=1in,vmargin=1in]{geometry} 
\usepackage[bottom]{footmisc}
\usepackage{fancyhdr}

\fancyhead{} % clear all header fields

\fancyhead[LO]{Samuel R. Bowman}
\fancyhead[RE]{Logical Reasoning in Compositional Vector Space Semantics}

\fancyfoot{} % clear all footer fields
\fancyfoot[CE, CO]{\thepage}

\renewcommand{\headrulewidth}{0.4pt}

\renewcommand{\footrulewidth}{0.4pt}

\pagestyle{fancy}
\setlength{\headheight}{15pt}

\usepackage{setspace}
\spacing{1}

\usepackage{enumitem}
\setlist{noitemsep}

\usepackage{float}

\usepackage[round]{natbib}
\bibpunct[,~]{(}{)}{,}{a}{,}{,}
\setlength{\bibsep}{0pt}

%----------------------------------------------------------------------------%
%                    Title, author, and affiliation                          %
%----------------------------------------------------------------------------%

\title{\Large Logical Reasoning in Compositional Vector Space Semantics}
\author{Samuel R. Bowman\\\textsc{Ling 236}}
%\\\\Qualifying Paper \textit{for}\\The Department of Linguistics, Stanford University\\\\Committee:\\Paul Kiparsky (chair), Arto Anttila, and Christopher Manning
% \date{Edinburgh P-Workshop, May 2013}
% \author{Max Bane and Jason Riggle}
% \date{draft -- \today}

%----------------------------------------------------------------------------%
%                    Fonts and encoding                                      %
%----------------------------------------------------------------------------%

\usepackage[dirty,tipa]{ucs}	% Utf-8
\usepackage[utf8x]{inputenc}	% fonts
\usepackage[T1]{fontenc}		% in Tex

\usepackage{tipa}\let\ipa\textipa
\usepackage{vowel}

\usepackage{pifont}
\newcommand{\BlankCell}{}

%\usepackage{palatino}           % body font 
\renewcommand{\ttdefault}{lmtt} % tt set to latinmodern

%----------------------------------------------------------------------------%
%                    Graphics, tables, and figures                           %
%----------------------------------------------------------------------------%
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{graphicx}
\usepackage{multido}
\usepackage{ot-tableau}
\usepackage{pstricks}
\usepackage{pst-slpe}

\psset{gridcolor=green, subgridcolor=white, dimen=middle,
	   linewidth=1pt, subgriddiv=1, gridlabels=4pt, gridwidth=0.2pt}
\psset{cornersize=absolute,linearc=0.06}

\newpsstyle{final}{linewidth=0.5pt,fillstyle=solid,fillcolor=gray95,%
				   doubleline=true,doublesep=0.8pt,dimen=outer}

%----------------------------------------------------------------------------%
%                    Algorithms and Theorems                                 %
%----------------------------------------------------------------------------%

\usepackage[ruled,noend,linesnumbered]{algorithm2e}

\usepackage[leqno,fleqn]{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{wasysym}

\usepackage{stmaryrd}

\newtheoremstyle%
{dotless}%      name
{\topsep}%      Space above
{\topsep}%      Space below
{\normalfont}%  Body font
{}%             Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}%    Thm head font
{}%             Punctuation after thm head
{\newline}%     Space after head: " " = interword space, \newline = linebreak
{}%             Thm head spec (can be left empty, meaning `normal')
%               alternative spec: {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}

\theoremstyle{dotless}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}

\newtheorem{exe}[thm]{Example}

%----------------------------------------------------------------------------%
%                    Hyperlinks                                              %
%----------------------------------------------------------------------------%
\newrgbcolor{lnkClr}{0.0 0.0 0.33}

\usepackage{hyperref}			% Clickable links in pdfs
\hypersetup{colorlinks,			% nice for large docs and 
			breaklinks,			% for links to web pages
            linkcolor=lnkClr,
			urlcolor=lnkClr,
            anchorcolor=lnkClr,
			citecolor=lnkClr}

%----------------------------------------------------------------------------%
%  Example environment
%----------------------------------------------------------------------------%

\newenvironment{example}
{\refstepcounter{equation}\begin{list}{(\theequation)}{\leftmargin=7mm \labelwidth=5mm}\item}%
{\end{list}}

\def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}

%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%%-%%%%%%%%
%./\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\..../\.%
%/  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \../  \%
%    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    \/    %
%%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%%% %%%%%%
\usepackage{gb4e}

\begin{document}
\maketitle
% \addtocounter{section}{-1}

\def\Y{\textbabygamma}
\def\P{\textceltpal}
\def\Rs{\textsubwedge} % restore this
\def\E{\tipaloweraccent[.1ex]{8}{e}} %restore this
\def\A{\sublptr{\"{a}}}

\maketitle

%TODO: See leichturm margins

%TODO: See Ken Shan slides

%TOOD: Mention Potts ministudy in 236 handout

%TODO: Replicate Baroni et al 2012

\section{Introduction}

Vector space models of semantic meaning have  substantially advanced the state of the art in applied natural language processing tasks like parsing, paraphrase detection, and sentiment analysis. These models have benefited in particular from two properties: The ability to learn representations of individual words that are \textit{optimized for the task at hand}, and their ability to compose those word representations into higher-order constituent representations \textit{in the same semantic vector space}. Together, these properties allow the model to capture nuanced compositional representations of sentence meaning without the need for any independent lexical semantic model.

There is not at this time any general understanding of what \textit{kind} of information these models are capable of learning and encoding in the vector space. It is clear that they are able to learn various useful facets of meaning well, but it is possible that many of the properties that are necessary for a full account of linguistic meaning like quantification and monotonicity are missed: while these models achieve state-of-the-art performance, the models that they supplant are largely models based on simple expert-designed features with substantial known limitations.

In this study, I propose to probe the representational capacities of these systems. In particular, I propose to build a model which learns task-specific compositional vector representations---in the style of Socher et al.---for the task of recognizing textual entailment in a regimented formal language. This model should shed light on the abilities of these schemes to capture reasoning patterns involving quantification, monotonicity, intersective and nonintersective modification, and other phenomena.

\section{The Pilot Experiment: Reasoning with idealized data}

If the pilot model works: Distributional system with X composition function gets quantification. Publish and push on how best to learn it in broader domain applications.

If it fails: Can we say that the system isn't powerful enough? Squib and return to even more supervised approaches.

\section{Background}

\citet{baroni2012entailment} present a real world version of the entailment reasoning problem, and have the current state of the art on that task. [Ken Shan is the last author on this.]

\begin{quote}
Interestingly, on the QN entailment task, neither our classifier trained on AN-N pairs nor the balAPinc method beat baseline methods. This suggests that our successful QN classifiers tap into vector properties beyond such relations as feature inclusion that those methods for nominal entailment rely upon.
\end{quote}

\begin{quote}
The FS definition of entailment has been modified
by taking common sense into account. Instead of
a relation from the truth of the consequent to the
truth of the antecedent in any circumstance, the
applied view looks at entailment in terms of plausibility \\\\... if a human who reads (and trusts) would most likely infer that is also true.
\end{quote}

\citet{erk2009representing} discusses notions of subsumption from distributional vectors. His results would be discouraging if we were trying to use pure distributional vectors rather than task-learned ones.

\begin{quote}
\textit{Baroni et al on Erk:} Erk (2009) suggests that it
may not be possible to induce lexical entailment
directly from a vector space representation, but it
is possible to encode the relation in this space after it has been derived through other means.\\\\
...run implies move in some contexts, but not all, for example not in the context computer runs.
\end{quote}

\citet{baroni2011we} present a corpus of word pairs manually annotated with relations like \textit{hypernym}, \textit{hyponym}, \textit{meronym}, \textit{associated event}, and \textit{common attribute}, as well as negative example relations.

\begin{quote}
\begin{verbatim}
missile-n	weapon	hyper	rocket-n
missile-n	weapon	hyper	vehicle-n
missile-n	weapon	hyper	weapon-n
missile-n	weapon	mero	engine-n
missile-n	weapon	mero	explosive-n
\end{verbatim}
\end{quote}

This may be useful as a fringe test case, but the lack of proper entailment/cover/equiv relationships may make it more difficult to test the logical capacity of the model by these means.

\citet{lenci2012identifying} have more of the same: distributional inclusion does capture some entailment information.

\begin{quote}
We introduce BLESS, a data set specifically
designed for the evaluation of distributional
semantic models. BLESS contains a set of tuples instantiating different, explicitly typed semantic relations, plus a number of controlled
random tuples. It is thus possible to assess the
ability of a model to detect truly related word
pairs, as well as to perform in-depth analyses of the types of semantic relations that a
model favors. We discuss the motivations for
BLESS, describe its construction and structure, and present examples of its usage in the
evaluation of distributional semantic models.
\end{quote}

In my notes for some reason: LDC's ACE08 Local Relation Detection and Recognition (LRDR) task

\citet{goller1996learning} present the core of the machine learning algorithm needed for the task-specific representations used in this project. I can't make heads or tails of their writing.

\citeapos{widdows2004geometry} book primarily serves as an primer on vector space semantics for those without sufficient mathematical background, but it does briefly address some aspects of the logical capacities of such systems. In particular, he presents a system for constructing information-retrieval targeted query vectors, and builds a practical quasi-logic around it which is capable of conjunction (\textit{rock} AND \textit{blues}) using vector addition, and subtraction (\textit{rock} NOT \textit{blues}) through the the operation of finding the orthogonal vector, expressed as:

\begin{equation}
a \text{~NOT~} b = a - (a \cdot b)b
\end{equation}

\citet{maccartney2009extended} \citep[from Bill's dissertation][]{maccartney2009natural} presents the framing of entailment and monotonicity used here, and shows that they can (with a lot of hacks) be used to do the Recognizing Textual Entailment task on real language. No vectors here.

\citet{icard2012inclusion} presents an extension to the NLI/MacCartney and Manning logic involving an elaborated notion of monotonicity. He uses the same seven basic entailment relations as MacCartney. The added detail relates to the model described in this paper as a formal description of what the model should ideally be able to learn, and not as a description of a formal property that should be externally introduced into the model.

\citet{socher2011semi} is the primary introduction to the idea of learning a set of quasi-distributional vectors and a composition function on those vectors, all guided to optimize performance on some task.

\citet{socher2011dynamic} extends this to paraphrase detection, an (easy, by our standards) task of comparing two learned sentence representations.

Socher et al. (in prep; ACL 2013) and \citet{chen2013learning} define a more powerful version of the composition function based on a third-order tensor. I'm using this layer as the top node that builds the comparison vector from which relation types are extracted.

Socher and co. have a few other recent and in-progress papers setting various task-performance records. Useful for motivation, but not that informative.

% Distributional Models of the Representation and Acquisition of Natural Language Categories, Lapata and Fountain: Likely worth ignoring

Basile et al: Encoding syntactic dependencies by vector permutation. Slides are a bit vague. Give it a closer look!

\section{Methods: The pilot model}

The initial model is built upon three functions: The composition function, which takes individual word representations and composes them into higher-order constituent representations, the comparison function, which takes pairs of word representations and compares them to yield a relation representation, and the softmax function, which turns this abstract relation representation into a probability distribution over the seven relation types. Each of these has numerous parameters which we learn from training data, and when composed appropriately, they form a relation classifier for pairs of sentences of arbitrary length.

In this section, I lay out the definitions of these three functions, how they are applied, and how their parameters can be learned.

\subsection{Forward computation: Identifying a single relation}

The figure shows an example of the structure over which information flows in a forward (classifying) run of the model. The leaves of the tree are the vector representations of individual words. The left input, ``dog'', is comprised of only a single word, and can be fed directly into the comparison function. ``very big cat'', however, contains three of these representations, which need to be reduced to a single vector. This is done by applying the composition function twice, yielding a vector of the same length as the words themselves at each application. The order of application of the composition function is determined by a constituency structure (unlabeled binary parse tree) provided with the input text. This structure is annotated manually for these inputs, but could in other cases be derived from the outputs of any of a large number of extant tree parsers.

Once the input has been reduced to two vectors of equal length, these are fed into the comparison layer, and subsequently into the softmax layer. The output of the softmax layer yields a distribution over possible relations between the two constituents, and choosing the maximum at this point yields the model's hypothesized relation.

\subsubsection{The comparison function}

The next layer is the comparison layer. This takes in the two $D$-dimensional representations of each constituent, $\vec{x^1}$ and $\vec{x^2}$, and transform them into a $D'$-dimensional vector $y$ representing the relation between the two. While this layer is not intended to be the same function as the composition function that builds up the individual constituents, I parameterize it in the same way, as an RNTN layer with three parameters, one a third-order tensor (in other words, a vector of matrices, marked $\mathbf{A^{[1...D']}}$), one a matrix  ($\mathbf{B}$), and one a vector ($\vec{c}$).

Each element in the output vector is computed using a single matrix drawn from the third order tensor parameter, a single row of the matrix parameter, and a single entry of the vector parameter, as follows.

\begin{equation}
y_i = f(\vec{x^1}^T \mathbf{A^{[1...D']}} \vec{x^2} + \vec{B_{i,:}} [\vec{x^1}; \vec{x^2}] + c_i)
\end{equation}

\begin{equation}
Cpr(\vec{x^1}, \vec{x^2}) = [y_1; y_2; ...; y_D']
\end{equation}

The result of each of these computations is fed into a sigmoid nonlinearity $f$. In my implementation, I use $f(x) = \tanh(x)$.

\subsubsection{The composition function}

As with the comparison layer, the composition layer is an RNTN layer with three parameters, one a third-order tensor ($\mathbf{K^{[1...D']}}$), one a matrix  ($\mathbf{L}$), and one a vector ($\vec{m}$). As above, the output of the layer is computed as follows.

\begin{equation}
p_i = f(\vec{x^1}^T \mathbf{K^{[1...D]}} \vec{x^2} + \vec{L_{i,:}} [\vec{x^1}; \vec{x^2}] + m_i)
\end{equation}

\begin{equation}
Mrg(\vec{x^1}, \vec{x^2}) = [y_1; y_2; ...; y_D']
\end{equation}

\subsubsection{The softmax layer}

TODO: Cite UFDL

The top layer of the model is a softmax classifier, which takes the model output and transforms it into a multinomial probability distribution over the seven possible relation types.

The softmax function takes the input from the layer below, denoted $\vec{y}$, appends a 1 to yield a bias term, multiplies the result by a matrix $\mathbf{S}$ and exponentiates it elementwise to yield an unnormalized probability measure.

\begin{equation}
\vec{\mu}(\vec{y}) = exp(\mathbf{S}\cdot[1; \vec{y}])
\end{equation}

From this, we can normalize to get a probability function by dividing each entry in the vector by the sum of the vector:
\begin{equation}
\vec{\pi}(\vec{y}) = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}
\end{equation}

\begin{equation}
P(rel(x_1, x_2)=r) = \pi(\vec{y})_r
\end{equation}

$\mathbf{S}$ above is a $D' \times R$ matrix.

\subsection{Backpropagation: Learning from a single relation}

The parameters of the model---both the function parameters and the word parameters---are learned using the L-BFGS algorithm TODO: CITE. L-BFGS is a TODO: CHECK

These gradients for each data point are computed using the backpropagation through structure algorithm \cite{goller1996learning}. In this algorithm, the model is first run forward in order to compute the predicted distribution, which is in turn used to compute the objective function.

\subsubsection{The Objective}

For an objective function, I use the negative logarithm of the probability of the true relation to capture correctness, and the sum of the squared parameter values (L2 normalization) to minimize overfitting. I present it here as a function of the two top-level vectors of the inputs, and the single correct relation specification.

\begin{equation}
E(\vec{w_1}, \vec{w_2}, \vec{r}, \theta) = - \sum_{i=1}^N \log(\pi(Cpr(\vec{w_1}, \vec{w_2}))_{r_i}) + \frac{\lambda}{2}\sum_j\theta_j^2
\end{equation}

Where $\theta$ is a single vector containing all of the entries of all of the model parameters:
\begin{equation}
\theta = \text{linearize}([\mathbf{A^{[1...D']}}, \mathbf{B}, \vec{c}, \mathbf{K^{[1...D]}}, \mathbf{L}, \vec{m}, \mathbf{S}])
\end{equation}

\subsection{Gradients for S}

The simplest gradient to compute is that of the last parameter to be applied, the matrix $\mathbf{S}$ in the probability function. Here I do so, assuming for simplicity that $N = 1$.

\begin{equation}
\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k} - \frac{\partial E}{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}}\frac{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}}{\partial S_{j,k}} = 
\end{equation}

Where $\pi(\vec{\mu}(\vec{y}))_{r_i}$ is the probability of the true relation.

\begin{equation}
\frac{\partial E}{\partial \pi(\vec{\mu}(\vec{y}))_{r_i}} = -\frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}}
\end{equation}

\begin{equation}
\vec{\mu}(\vec{y}) = exp(\mathbf{S}\cdot[1; \vec{y}])
\end{equation}

\begin{equation}
\vec{\pi}(\vec{y})_r = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}
\end{equation}

if $r \neq j$

%f = \vec{\mu}(\vec{y})_r
%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
%f' = 0
%g' = [1; \vec{y}]_{k}
%
%if r \neq j

\begin{equation}
\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{-\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
\end{equation}

\begin{equation}
\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k} - \frac{ }{\pi(\vec{y})_{r_i}} \frac{\vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
\end{equation}

%f = \vec{\mu}(\vec{y})_r
%g = \sum_{r' = 1}^R \mu(\vec{y})_{r'}
%f' = [1; \vec{y}]_{k}
%g' = [1; \vec{y}]_{k}
%

if eq

\begin{equation}
\frac{\partial \pi(\vec{y})_{r_i}}{\partial S_{j,k}} = \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
\end{equation}

\begin{equation}
\frac{\partial E}{\partial S_{j,k}} = \lambda S_{j,k}  + \frac{ 1 }{\pi(\vec{\mu}(\vec{y}))_{r_i}} \frac{[1; \vec{y}]_{k}\cdot \sum_{r' = 1}^R \mu(\vec{y})_{r'} - \vec{\mu}(\vec{y})_r\cdot [1; \vec{y}]_{k}}{(\sum_{r' = 1}^R \mu(\vec{y})_{r'})^2}
\end{equation}

%d pi_ri  / d mu

%f = \vec{\mu}(\vec{y})_r
%g =

%\vec{\pi}(\vec{y}) = \frac{\vec{\mu}(\vec{y})_r}{\sum_{r' = 1}^R \mu(\vec{y})_{r'}}

%\log(\pi(Cpr(\vec{w_1}, \vec{w_2}))_{r_i})

\subsection{The comparison function}

Since we use $\tanh(x)$ for the sigmoid function $f$, we have that $\frac{df}{dx} = sech^2(x)$.

\bibliography{../../../Bibs/MLSemantics}{}
\bibliographystyle{plainnat}

\end{document}