### 2014-06-03 writing/S'14 paper/discussion.tex
Updating paper.

\section{General discussion}

These experiments represent one of the first attempts to reproduce any large fragment of the behavior of a complex logic within a neural network model, and the first attempt that we are aware of to address either the encoding of lexical relations or the learning of recursive operators. This presents considerable challenges in evaluating the particular models that we choose, since we cannot rely on prior results to establish that any particular amount or type of training data is sufficient to teach any model the structure of the logic. The positive results that we have fonud, however, are extremely promising for the future of learned representation models in the applied modeling of meaning. We have seen that recursive neural tensor networks are able to encode lexical relations accurately and encode recursive operators. We have also seen that both RNNs and RNTNs are able to handle the meanings of quantifiers in an inference setting in at least some cases.

There is ample room to build on these results. In the interest of fully mirroring the capacity of existing natural logics in learned models, it would be valuable to extend these experiments to cover other ways in which meanings are encoded in natural language, including challenges such as reasoning over sentences with transitive verbs or relative clauses. In addition, it would be highly informative to compare these results on standard recursive neural networks with other proposed learned models for sentence meaning, such as dependency tree RNNs \cite{sochergrounded}, Belief Propagation RNNs (TODO: cite), or convolutional RNNs \cite{kalchbrenner2014convolutional}.

% TODO: One para of closing discussion.
% - Results strongly suggest that RNTNs can encode the basics of this logic. First such result.
% -- Decay over size in recursion results not ideal, but this kind of behavior may not be problematic since there is a practical bound on the length of natural language sentences.
% - Future work: 
% -- Harder problems
% --- Sentences with more types of structure: Transitive verbs, relative clauses, etc.
% -- More models
% --- BP-RNNs, conv-RNNs