### 2014-06-04 writing/S'14 paper/recursion.tex
Intermediate paper update.

\section{Recursive structure}\label{sec:recursion}

% TODO: Cite Chomsky/Hauser/Fitch? -- no, not appropriate here --- CP.

Recursive structure is an prominent feature of natural languages and
an important part of their expressivity, as it allows people to use
and interpret complex structures they have never encountered before.
Theoretical accounts of natural language syntax and semantics
therefore rely heavily on recursive structures, and we expect that
accurate computational models will have to come to grips with them as
well. The present section pursues this question for our two classes of
RNN. In evaluating the model, we take advantage of the fact that
recursive structures of this kind define potentially infinite
languages, by testing the model on strings that are longer and more
complex than any seen in testing.

% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
%   said hello}, and \ii{Carl thinks that Bob said that Alice said
%   hello}. Overt recursion of this kind is easy to find, and
% theoretical accounts of natural language syntax and semantics rely
% heavily on recursive structures. In order for a model to be able to
% accurately learn natural language meanings, then, we expect that it
% would need to be able to learn to represent the meanings of function
% words in a such a way that they are able to behave correctly when
% taking their own outputs as input. In evaluating the model, we take
% advantage of the fact that recursive structures of this kind define
% potentially infinite languages by testing the model on strings that
% are longer and more complex than any seen in testing.

As in Section~\ref{sec:join}, we test this phenomenon within the
framework of natural logic, but we now replace the unanalyzed symbols
from that experiment with expressions that involve recursive
structure. These expressions represent a truth-functionally complete classical
propositional logic: each atomic variable denotes either $\True$ or
$\False$, and the only operators are truth-functional ones.
Table~\ref{tab:pl} defines this logic, and Table~\ref{tab:plexs} gives
some examples of relational statements that we can make
in these terms. To compute these relations between statements, we
exhaustively enumerate the sets of assignments of truth values to
proposition variables that would satisfy each of the statements and
then convert the set-theoretic relation between those assignments into
one of the seven relations in Table~\ref{b-table}. As a result, each relational
statement represents a valid theorem of the propositional logic.

\begin{table}[htp]
  \centering
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $a$, $b$, $c$, $d$, $e$, $f$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \quad
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg a$        & $\natneg$ & $a$ \\
      $\plneg \plneg a$ & $\nateq$  & $a$ \\
      $a$               & $\natfor$ & $(a \plor b)$ \\
      $a$               & $\natrev$ & $(a \pland b)$ \\
      $(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg(\plneg a \pland \plneg b)$ & $\nateq$ & $(a \plor b)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of statements about relations between
      well-formed formulae, defined in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Propositional logic with natural logic relations.}  
  \label{prop-figure}
\end{table}

% NOTE: There was some confusion at CSLI and the 224U lecture about
% whether this was theorem proving or some kind of model checking.
% Hopefully this is explicit enough.
Socher et al.~\cite{socher2012semantic} demonstrate the learning of a
logic in a matrix-vector RNN model somewhat similar to our own, but
the logic discussed here is substantially stronger, and a much better
approximation of the kind of structure that is needed for natural
language. The logic learned in that experiment is boolean, wherein the
atomic symbols are simply the values $0$ and $1$, rather than
variables over those values. While learning the operators of that
logic is not trivial, the outputs of each operator can be represented
accurately by a single bit. The statements of propositional logic
learned here describe conditions on the truth values of propositions
where those truth values are not known. As opposed to the two-way
contrasts seen in \cite{socher2012semantic}, this logic distinguishes
between $2^{6} = 64$ possible assignments of truth values, and
expressions of this logic define arbitrary conditions on these
possible assignments, for a total of $2^{64}$ %($\approx 10^{20}$)
possible statements that the intermediate vector representations need 
to be able to distinguish.

For our experiments, we randomly generate a large set of  unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is a tautology or
contradiction (a statement that is true of either all or no possible
assignments), for which none of the seven relation labels in
Table~\ref{b-table} can hold. The resulting set of formula pairs is
then partitioned into bins based on the number of logical operators in
the longer of the two formulae. We then randomly sample 15\% of each
bin for a held out test set.

If we do not implement any constraint that the two statements
being compared are similar in any way, the generated data consists in
large part of statements in which the two formulae refer to largely
separate subsets of the six variables, and to which we will nearly
always assign the $\natind$ relation. In an effort to balance the
distribution of relation labels without departing from the basic task
of modeling propositional logic, we disallow individual pairs of
statements from referring to more than four of the six proposition
variables.

We trained both the RNN and RNTN models on the data of size four or
less (65k pairs), and tested it on examples of up to size 12 (44k
pairs). We initialized the model parameters randomly, including the
vector representations of the six variables. The results are shown in
Figure~\ref{prop-results}. In tuning, we found that the RNN model was
approximately optimal with 45-dimensional vector representations, and
the RNTN model was approximately optimal with 25 dimensions. We fixed
the size of the feature vector for the classifier at 75 dimensions. We
found that the RNTN model was able to perform almost perfectly on
unseen small test examples, with accuracy above 99\% below size four.
After depth four, performance gradually falls with increasing size.
The RNN model did not perform well, reaching only 88.2\% accuracy on
the smallest test examples, and declining from there to near-baseline
performance at size 12.

\begin{figure}[htp]
  \centering
  \includegraphics[width=5.5in]{recursion\string_results.eps}
  \caption{Model performance on propositional logic, by expression size.}  
  \label{prop-results}
\end{figure}

The performance of the RNTN model on small unseen test examples
indicates that it learned a correct approximation of the underlying
logic. It appears that this approximation is accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
depth, resulting in the gradually dropping performance. This is not
necessarily a significant flaw in the model, since it remains possible
that a less lossy approximation of the same logic would do no better
on the training objective, since that objective only considers short
strings and includes a regularization term. Whether a more accurate
encoding could be learned from a larger set of training data remains a
quention for future work. In contrast, the RNN model did not appear to
be able to learn a correct approximation of the logic.