### 2014-06-04 writing/S'14 paper/quantifiers.tex
Adding new Figure 1, updating results on quantifiers.

\section{Reasoning with natural language quantifiers}\label{sec:quantifiers}

Even to the extent that RNTN models can handle functional meanings in
the form of the operators of propositional logic, this is not a
guarantee that these models can handle functional meanings of the form
seen in natural language. As a first step towards investigating this
issue, we attempt to directly measure the degree to which RNNs are
able to develop suitable representations for the semantics of natural language
quantifiers like \ii{some} and \ii{all}. Quantification is far from
the only place in natural language where complex functional meanings
are found, but it is a natural starting point, since it can be tested
in sentences whose structures are otherwise quite simple, and since it
has formed a standard case study in prior formal work on natural
language inference.

% \subsection{Data}

This experiment replicates similar work described in
\cite{bowman2013can}, which found that RNTNs can learn to reason well
with quantifier meanings given sufficient training data. This paper
replaces the partially manually annotated data in that paper with data
that is generated directly using the logical system that we hope to
model, yielding results that we believe to be substantially more
straightforward to interpret.

Our data consists of pairs of sentences generated from a small
artificial grammar. Each sentence contains a quantifier, a noun which
may be negated, and an intransitive verb which may be negated. We use
the basic quantifiers \ii{some}, \ii{most}, \ii{all}, \ii{two}, and
\ii{three}, and their negations \ii{no},
\ii{not-all}, \ii{not-most}, \ii{less-than-two}, and
\ii{less-than-three}. We also include five nouns, four intransitive
verbs, and the negation symbol \ii{not}. In order to be able to define
relations between sentences with differing lexical items, we define
the lexical relations between each noun--noun pair, each verb--verb
pair, and each quantifier--quantifier pair. The grammar accepts
aligned pairs of sentences of this form and calculates the natural
logic relationship between them.  Some examples of these data are
provided in Table~\ref{examplesofdata}.  As in previous sections, the
goal of learning is then to assign these relational labels accurately
to unseen pairs of sentences.

%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
%verbs = ['walk', 'move', 'swim', 'growl']
%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
%adverbs = ['', 'not']

% To assign relation labels to sentence pairs, we built a small
% task-specific implemenation of MacCartney's logic that can
% accurately label sentences of this restricted language. The logic is
% not able to derive all intuitively true relations of this language,
% and fails to derive a single unique relation for certain types of
% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
% $\natneg$ (some pet) (not growl)}), and we simply discard these
% examples. Exhaustively generating the valid sentences under this
% grammar and choosing those to which a relation label can be assigned
% yields 66k sentence pairs. Some examples of these data are provided
% in Table~\ref{examplesofdata}.

\begin{table}[htp]
  \centering
  \begin{tabular}{l c l}
    \toprule
    (most warthogs) walk         & $\natneg$ & (not-most warthogs) walk\\
    (most mammals) move          & $\natind$ &  (not-most (not turtles)) move\\
    (most (not pets)) (not swim) & $\natrev$ & (not-most (not pets)) move 
    \\[2ex]    
    (no turtles) (not growl)     & $\natalt$ & (no turtles) (not swim)\\
    (no warthogs) swim           & $\natrev$ & (no warthogs) move\\
    (no warthogs) move           & $\natfor$ & (no (not reptiles)) swim\\
    \bottomrule
  \end{tabular}
  \caption{Sample data involving two different quantifier pairs.}
  \label{examplesofdata}
\end{table}

We evaluate the model using two experimental settings. In the simpler
setting, \textsc{all split}, we randomly sample 85\% of the data and evaluate on the
remaining 15\%. In this setting, the model is being asked to learn a
complete reasoning system for the limited language and logic presented
in the training data, but it is not being asked to generalize to test
examples that are substantially different from those it was trained
on. Crucially though, to succeed on this task, the model must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
\eqref{p1} and \eqref{p2} in training and, from that information,
determine \eqref{p3}.
%
\begin{gather}
  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
  \\
  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
  \\
  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
\end{gather}

While our primary interest is in discovering the extent to which our
models can learn to encode the logic given an arbitrary amount of
data, we are also interested in the degree to which they can infer a
correct representation for the logic from more constrained training
data. To this end, we segment the sentence pairs according to which
quantifiers appear in each pair, and then hold out one such pair for
testing. We hypothesize that a model that can efficiently learn to
represent a logic should be able to construct an accurate
representation of each held out quantifier from the way that it
interacts with the other nine quantifiers which are not held
out. Since running this experiment requires choosing a pair of
quantifiers to hold out before training, the resource demands of
training prevent us from testing each of the 55 possible possible
pairs of quantifiers, and we choose only four pairs to test on.  Three
of these (\ii{two}/\ii{less-than-two}, \ii{not-all}/\ii{not-most}, and
\ii{all}/\ii{some}) were chosen because they provide allow for the most
different class labels at in the training data. The fourth is a
self-pair (\ii{no}/\ii{no}), meant to test that the model correctly
handles equality.

% Further explanation for motivation for no/no?

\begin{table}[htp]
  \centering
  \begin{tabular}{ l rrr }
    \toprule
    Data & Most frequent class & TODO dim RNN (\%) & 16 dim RNTN (\%)\\
    \midrule
    \textsc{all split}	& 35.4 &	63.9&	\textbf{93.4}
    \\[1ex]    
    \textsc{pair two/less-than-two}	& 59.8 &	79.5 &	\textbf{98.3} \\
    \textsc{pair not-all/not-most}	&0 &	   62.2  &	\textbf{83.3} \\
    \textsc{pair all/some}	& 0& 63.6  &	\textbf{81.0} \\
    \textsc{pair no/no}	& 30.8 &	66.0 &	\textbf{100} \\
    \bottomrule
  \end{tabular}
  \caption{Quantifier experiment performance. }
  \label{resultstable}
\end{table}

The results for these experiments are shown at the top of Figure \ref{resultstable}. We compare the results to a most frequent class baseline, which reflects the frequency in the test data of most frequent class in the training data, $\natind$.

\mynote{
The RNN model was approximately optimal with N dimensional word representations and an M dimensional comparison layer. The RNTN was approximately optimal with N and M dimensions, respectively.
% TODO: Update dimensionality
(Note: Models still running. Results might improve and this para is
likely to change.)

The perfect performance by the RNTN on the
\textsc{pair no/no} experiment and its strong performance on
\textsc{pair two/less-than-two} suggests that it is at least plausible
that this model is able to handle quantifiers correctly given
sufficient training data, but the sub 90\% results on two of the
training settings suggest that the model may not be able to
generalize from data this impoverished in general. However, the fact
that both models perform far above baseline is promising, and the
question of how much data is necessary to accurately capture
quantifier behavior in a na\"ive model remains open.

Discussion of RNN results (and updated discussion of better RNTN results) to be filled in as results arrive.\\...\\... % TODO
}