### 2014-08-25 writing/BayLearn abstract/long_abstract.tex
Updated draft abstract.

\documentclass{article} % For LaTeX2e
\usepackage{nips14submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage[leqno, fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{qtree}
\usepackage[numbers]{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xcolor}

\definecolor{mylinkcolor}{rgb}{0,0,0} % black
\hypersetup{colorlinks, linkcolor=mylinkcolor, urlcolor=mylinkcolor, citecolor=mylinkcolor}

\newcommand{\nateq}{\equiv}
\newcommand{\natind}{\mathbin{\#}}
%\newcommand{\natneg}{\raisebox{2px}{\tiny\thinspace$\wedge$\thinspace}}
\newcommand{\natneg}{\mathbin{^{\wedge}}}
\newcommand{\natfor}{\sqsubset}
\newcommand{\natrev}{\sqsupset}
\newcommand{\natalt}{\mathbin{|}}
\newcommand{\natcov}{\mathbin{\smallsmile}}

\newcommand{\plneg}{\mathop{\textit{not}}}
\newcommand{\pland}{\mathbin{\textit{and}}}
\newcommand{\plor}{\mathbin{\textit{or}}}

% Strikeout
\newlength{\howlong}\newcommand{\strikeout}[1]{\settowidth{\howlong}{#1}#1\unitlength0.5ex%
\begin{picture}(0,0)\put(0,1){\line(-1,0){\howlong\divide\unitlength}}\end{picture}}

\newcommand{\True}{\texttt{T}}
\newcommand{\False}{\texttt{F}}
\usepackage{stmaryrd}
\newcommand{\sem}[1]{\ensuremath{\llbracket#1\rrbracket}}

\renewcommand{\bibsection}{\subsubsection*{References}}

\usepackage{gb4e}

\def\ii#1{\textit{#1}}

\newcommand{\mynote}[1]{{\color{red}\framebox{\begin{tabular}{p{0.9\textwidth}}\footnotesize#1 \end{tabular}}}}

\title{Recursive Neural Networks for Learning Logical Semantics}

\author{
Samuel R.\ Bowman$^{\ast\dag}$ \\
\texttt{sbowman@stanford.edu} \\[2ex]
$^{\ast}$Stanford Linguistics \\
\And
Christopher Potts$^{\ast}$\\
\texttt{cgpotts@stanford.edu} \\[2ex]
$^{\dag}$Stanford NLP Group
\And
Christopher D.\ Manning$^{\ast\dag\ddag}$\\
\texttt{manning@stanford.edu}\\[2ex]
$^{\ddag}$Stanford Computer Science
}

% \author{
% Samuel R.\ Bowman \\
% NLP Group, Dept.\ of Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{sbowman@stanford.edu}
%  \And
%  Christopher Potts \\
% Dept.\ of Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{cgpotts@stanford.edu}
%  \And
% Christopher D.\ Manning \\
% NLP Group,  Depts.\ of Computer Science and Linguistics\\
% Stanford University\\
% Stanford, CA 94305-2150 \\
% \texttt{manning@stanford.edu}
% }

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
\maketitle

% Compare with Grefenstette etc?

  Supervised recursive neural network models (RNNs) for sentence
  meaning have been successful in an array of sophisticated language
  tasks, but it remains an open question whether they can learn
  compositional semantic grammars that support logical deduction.  We
  address this question directly by for the first time evaluating
  whether each of two classes of neural model --- plain RNNs and
  recursive neural tensor networks (RNTNs) --- can correctly learn
  relationships such as entailment and contradiction between pairs of
  sentences, where we have generated controlled data sets of sentences
  from a logical grammar.  Our first experiment evaluates whether
  these models can learn the basic algebra of logical relations
  involved. Our second and third experiments extend this evaluation to
  complex recursive structures and sentences involving quantification.
  We find that the plain RNN achieves only mixed results on all three
  experiments, whereas the stronger RNTN model generalizes well in
  every setting and appears capable of learning suitable
  representations for natural language logical inference.

\subsection*{Recursive neural network models}

\begin{figure}[hp]
  \centering\resizebox{4.5in}{!}{
  \input{figure1}}
  % \caption{The model structure used to compare \ii{((all reptiles) walk)} and \ii{((some turtles) move)}. 
  %  The same structure is used for both the RNN and RNTN layer functions.} 
  \label{sample-figure}
\end{figure}

In our experiments, we train pairs of recursive (tree structured) neural network models \cite{socher2013acl1} which are joined together with a shared top layer that generates features for a classifier. The classifier predicts the logical relation that holds between the sentences represented by the two trees. For an activation function, we use either a plain NN layer or a tensor combination layer.

% TODO gold parse structures

\subsection*{Inference and the semantic relations}

Our aim is to evaluate the ability of learned recusive models to represent semantic structure. We pursue this through an inference task, where the models must learn to choose the logical relation that applies between a pair of statements. The possible relations are the seven below from \cite{maccartney2009extended}. This is distinct from the limited prior work on learning neural network models for formal semantics \cite{grefenstette2013towards, rocktaschellow}, which use an interpretation task in which the model evaluates sentences as \textit{true} or \textit{false} with respect to some representation of the world. Our approach allows us to sidestep serious open problems involved in the representation of a complete set of knowledge about the world.

\begin{table}[h]\small
\centering\resizebox{4.5in}{!}{
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symbol & Set-theoretic definition & Example \\ 
    \midrule
    entailment         & $x \natfor y$   & $x \subset y$ & \ii{turtle, reptile}  \\ 
    reverse entailment & $x \natrev y$   & $x \supset y$ & \ii{reptile, turtle}  \\ 
    equivalence        & $x \nateq y$    & $x = y$       & \ii{couch, sofa} \\ 
    alternation        & $x \natalt y$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ & \ii{turtle, warthog} \\ 
    negation           & $x \natneg y$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$    & \ii{able, unable} \\
    cover              & $x \natcov y$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ & \ii{animal, non-turtle} \\ 
    independence       & $x \natind y$   & (else) & \ii{turtle, pet}\\
    \bottomrule
  \end{tabular}}
  \label{b-table}
\end{table}

\subsection*{Reasoning with atomic symbols}

If any model is to learn the behavior of a relational logic like the one presented here from a finite amount of data, it must learn to deduce new relations from already seen relations. Our first experiment evaluates the ability of our models to do this over pairs of atomic symbols. The model is trained on a randomly generated set of consistent statements like \{$a \natfor b$, $b \natneg c$, $c \nateq d$\}, and tested on novel examples that follow from the statements seen in training, like \{$b \natneg d$\}. A tuned RNTN reached greater than 99\% test accuracy, but no plain RNN surpassed 89\%.

\subsection*{Recursive structure in propositional logic}\label{sec:recursion}

Our second experiment introduces compositional structure to our examples, replacing the atomic statements above with pairs of short statements of propositional logic, like $\plneg a\natalt((a~(\pland~ b))~(\pland~c))$. We train our models on only pairs of statements with up to four symbols, but observe that the RNTN performs reasonably both on pairs of that length and on much longer test pairs, as shown below.

\begin{figure}[h]
  \centering
  \includegraphics[width=4in]{recursion\string_results\string_final.eps}
  \label{prop-results}
\end{figure}

\subsection*{Reasoning with natural language quantifiers and negation}\label{sec:quantifiers}

For our third experiment, we generate pairs of sentences in which each sentence contains one quantifier, and any of a small set of common nouns, as in the example \textit{(no warthogs) move $\natfor$ (no (not reptiles)) swim}. The parentheses indicate the tree structure for each sentence as it will be used by the model. We defined several different types of train--test split for this experiment. RNTNs performed either well ($>85\%$ accuracy) or perfectly on all of them, while plain RNNs did not break 80\% in any setting.

\bibliographystyle{unsrtnat}

\small % Note: Explicitly allowed in style guide
\bibliography{MLSemantics}

\end{document}