### 2014-10-09 writing/AAAI SSS '15/methods.tex

\subsection*{Neural network models for relation classification} \label{methods}

% TODO: Should we cite our NIPS manuscript?

We build a model to learn embeddings using the method of
\citet{Bowman:Potts:Manning:2014}, which is centered on the problem of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture of the model that we use, which is
limited to only pairs of single symbols (such as words), is depicted
in Figure~\ref{sample-figure}. The model represents the two input
symbols as embeddings, which are fed into a comparison function based
on one of two types of neural network layer functions to produce a
representation for the relationship between the two symbols. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, are trained through backpropagation with AdaGrad
\cite{duchi2011adaptive}.

\begin{figure}[tp]
  \centering
  \input{figure1}
  \caption{The model structure used to compare \ii{turtle} and \ii{animal}.} 
  \label{sample-figure}
\end{figure}

For a comparison funtion, we evaluate versions of the model with both
a plain NN and the NTN layer function proposed in
\citet{chen2013learning}, which adds a third order tensor parameter to
mediate multiplicative interactions between the two inputs. A leaky
ReLU nonlinearity \cite{maasrectifier} is applied to the output of
either layer function.
%
\begin{gather} \label{rnn}
\vec{y}_{\textit{NN}} = f(\mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b}) ~~~~~~~~~~~~~~~ \vec{y}_{\textit{NTN}} = f(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)} + \mathbf{M} [\vec{x}^{(l)}; \vec{x}^{(r)}] + \vec{b})
\end{gather} 
%

This model differs from that of \citet{Bowman:Potts:Manning:2014} in
two ways. Because the inputs are single symbols, there is no need for
the composition functions used in that (and prior) work. Also, for our
experiment on WordNet data, we introduce an additional neural network
layer between the embedding input and the comparison function, which
facilitates initializing the embeddings themselves from pretrained
vectors and was found to help peformance in that setting.

%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?