### 2014-10-09 writing/AAAI SSS '15/methods.tex
One more methods tweak.

\subsection*{Neural network models for relation classification} \label{methods}

We build an embedding based model using the method of
\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture of the model that we use, which is
limited to only pairs of single symbols (such as words), is depicted
in Figure~\ref{sample-figure}. The model represents the two input
symbols as embeddings, which are fed into a comparison function based
on one of two types of neural network layer functions to produce a
representation for the relationship between the two symbols. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, are trained using backpropagation.

The simpler version of the comparison simply concatenates the two input
vectors before feeding them into a standard neural network (NN) layer.
The more powerful neural tensor network (NTN) version uses an additional 
third-order tensor parameter to allow for multiplicative
interactions between the two inputs. For more details on the implementation and
training of the layer functions, see \cite{Bowman:Potts:Manning:2014}.

\begin{figure}[tp]
  \centering
  \input{figure1}
  \caption{The model structure used to compare \ii{turtle} and \ii{animal}.} 
  \label{sample-figure}
\end{figure}

This model used here differs from the one described in that work in
two ways. Because the inputs are single symbols, we do not use
a composition function here. Also, for our experiment on WordNet data, 
we introduce an additional neural network layer between the embedding 
input and the comparison function, which facilitates initializing the 
embeddings themselves from pretrained vectors and was found to help 
peformance in that setting.

%\ii{Source code and generated data will be released after the conclusion of the review period.} % TODO: Or upon request? Attach anonymized code?