### 2015-02-24 writing/F'14 paper/quantifiers.tex

\section{Reasoning with quantifiers and negation}\label{sec:quantifiers}

We have seen that recursive models can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a key test of whether our models 
can capture this complexity, we now directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
and negation are far from the only place in natural language where complex functional meanings
are found, but they are natural focus, since they have
formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}..

\paragraph{Experiments}
Our experimental data consist of pairs of sentences generated from a
small artificial grammar. Each sentence contains a quantifier, a noun
which may be negated, and an intransitive verb which may be
negated. We use the basic quantifiers \ii{some}, \ii{most}, \ii{all},
\ii{two}, and \ii{three}, and their negations \ii{no}, \ii{not-all},
\ii{not-most}, \ii{less-than-two}, and \ii{less-than-three}. We also
include five nouns, four intransitive verbs, and the negation symbol
\ii{not}. In order to be able to define relations between sentences
with differing lexical items, we define the lexical relations between
each noun--noun pair, each verb--verb pair, and each
quantifier--quantifier pair. The grammar accepts aligned pairs of
sentences of this form and calculates the natural logic relationship
between them.  Some examples of these data are provided in (\ref{p1}--\ref{p3}) below.
  As in previous sections, the goal of
learning is then to assign these relational labels accurately to
unseen pairs of sentences.

%nouns = ['warthogs', 'turtles', 'mammals', 'reptiles', 'pets']
%verbs = ['walk', 'move', 'swim', 'growl']
%dets = ['all', 'not_all', 'some', 'no', 'most', 'not_most', 'two', 'lt_two', 'three', 'lt_three']
%adverbs = ['', 'not']

% To assign relation labels to sentence pairs, we built a small
% task-specific implemenation of MacCartney's logic that can
% accurately label sentences of this restricted language. The logic is
% not able to derive all intuitively true relations of this language,
% and fails to derive a single unique relation for certain types of
% statement, including De Morgean's laws (e.g. \ii{(all pets) growl
% $\natneg$ (some pet) (not growl)}), and we simply discard these
% examples. Exhaustively generating the valid sentences under this
% grammar and choosing those to which a relation label can be assigned
% yields 66k sentence pairs. Some examples of these data are provided
% in Table~\ref{examplesofdata}.

In each run, we sample 20\% of the 60k generated pairs to test on and 
train on the remainder. In this setting, each model must learn a
complete reasoning system for the limited language and logic presented
in the training data, and must be able
to recognize all of the lexical relations between the nouns, verbs,
and quantifiers and how they interact. For instance, it might see
pairs like \eqref{p1} and \eqref{p2} in training and be required to 
then label \eqref{p3}.

\vspace{-0.6cm}
\begin{gather}
  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
  \\
  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
  \\
  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
\end{gather}\vspace{-0.8cm}

In addition to the two models under study, we also train a summing neural network as a baseline, 
which is identical to the RNN, except that instead of learning a parametric composition function,
it simply sums the word vectors in a sentence to compose them. This baseline model does not
use word order information, though the highly consistent sentence structure in this experiment means
that it can potentially distinguish any two valid sentences except in case the two differ only in the placement of the word \ii{not}, potentially introducing errors into the 58.8\% of examples that contain a sentence with a single
instance of \ii{not}.

\begin{table}[tp]
  \centering
  \begin{tabular}{ l r@{ \ }r r@{ \ }r r@{ \ }r }
    \toprule
    ~ & \multicolumn{2}{c}{25d SumNN} & \multicolumn{2}{c}{25d RNN}  & \multicolumn{2}{c}{25d RNTN}\\
    \midrule
    Train &	96...&	(95...)&	\textbf{100}&		\textbf{(100)}&	99.5&	(99.3)\\	
    Test & 90...&	(80...)&	\textbf{99.4}&\textbf{(99.5)}& \textbf{99.4} & (99.2)\\
    \bottomrule
  \end{tabular}
  \caption{Performance on the quantifier experiments, given as \% correct and macroaveraged F1. The most frequent class, $\natind$, appears in 35.4\% of examples.}
  \label{qresultstable}
\end{table} 
% TODO: Fill in baseline results

%
% do not allow a blank line --- adds too much space
%
\paragraph{Results} The results, reported in Table~\ref{qresultstable}, show that both recursive models are able to learn to generalize the underlying logic satisfactorily, and that the baseline model generalizes well but far from perfectly. We did not observe any consistent pattern in the few errors made by either model, and errors did not tend to be consistent across runs of the experiment, suggesting that there is no fundamental obstacle to learning a perfect model for this problem.

% TODO: Baseline