### 2015-02-25 writing/F'14 paper/recursion.tex

\section{Recursive structure}\label{sec:recursion}
% TODO: Enter baseline results, use [[Tree]]RNN.

% TODO: Clean up next sent.
Recursive structure is a prominent feature of natural language. 
This section examines the degree to which our models are able to learn a 
compositional semantics over recursive structures.
In our evaluations, we exploit the fact that our logical language
is infinite by testing on strings that are longer and more complex
than any seen in training.

% Consider, for example, \ii{Alice said hello}, \ii{Bob said that Alice
%   said hello}, and \ii{Carl thinks that Bob said that Alice said
%   hello}. Overt recursion of this kind is easy to find, and
% theoretical accounts of natural language syntax and semantics rely
% heavily on recursive structures. In order for a model to be able to
% accurately learn natural language meanings, then, we expect that it
% would need to be able to learn to represent the meanings of function
% words in a such a way that they are able to behave correctly when
% taking their own outputs as input. In evaluating the model, we take
% advantage of the fact that recursive structures of this kind define
% potentially infinite languages by testing the model on strings that
% are longer and more complex than any seen in testing.

\paragraph{Experiments}
As in \S\ref{sec:join}, we test this phenomenon on data generated from a logic,
 but we now replace the unanalyzed symbols
from that experiment with complex formulae. These formulae
represent a truth-functionally complete classical propositional logic:
each atomic symbol is a variable over the domain \{$\True$, $\False$\}, and the only
operators are truth-functional ones.  Table~\ref{tab:pl} defines this
logic, and Table~\ref{tab:plexs} gives examples of the some of the shorter relational
statements that we generate from it as train and test data. To compute these relations
between statements, we exhaustively enumerate the sets of assignments
of truth values to propositional variables that would satisfy each of
the statements, and then we convert the set-theoretic relation between
those assignments into one of the seven relations in
Table~\ref{b-table}. As a result, each relational statement represents
a valid theorem of the propositional logic, and and to succeed, the models 
must learn to reproduce the behavior of a theorem prover.\footnote{
Socher et al.~\shortcite{socher2012semantic} show that a matrix-vector TreeRNN
model somewhat similar to our TreeRNTN can learn boolean logic, 
a logic where the atomic symbols are simply the
values $\True$ and $\False$. While learning the operators of that logic is not trivial, the outputs of
each operator can be represented accurately by a single bit.
In the more demanding task presented here, the atomic symbols are variables over these values, and the sentence vectors must thus be able to distinguish up to $2^{64}$ distinct conditions on valuations.}

\begin{table}[tp]
  \centering\small
  \begin{subtable}[t]{0.45\textwidth}
    \centering
    \begin{tabular}[t]{l l}
      \toprule
      Formula     & Interpretation \\
      \midrule
      $p_1$, $p_2$, $p_3$, $p_4$, $p_5$, $p_6$ & $\sem{x} \in \{\True, \False\}$ \\
      $\plneg \varphi$ & $\True$ iff $\sem{\varphi} = \False$ \\
      $(\varphi \pland \psi)$ & $\True$ iff $\False \notin \{\sem{\varphi}, \sem{\psi}\}$ \\
      $(\varphi \plor \psi)$  & $\True$ iff $\True \in \{\sem{\varphi}, \sem{\psi}\}$ \\
      \bottomrule
    \end{tabular}    
    \caption{Well-formed formulae. $\varphi$ and $\psi$
      range over all well-formed formulae, and $\sem{\cdot}$ is
      the interpretation function mapping formulae into $\{\True,
      \False\}$.}\label{tab:pl}
  \end{subtable}
  \begin{subtable}[t]{0.45\textwidth}
    \centering\vspace{4mm}
    \begin{tabular}[t]{r c l}
      \toprule
      $\plneg p_3$        & $\natneg$ & $p_3$ \\
      $\plneg \plneg p_6$ & $\nateq$  & $p_6$ \\
      $p_3$               & $\natfor$ & $(p_3 \plor p_2)$ \\
      $(p_1 \plor (p_2 \plor p_4))$               & $\natrev$ & $(p_2 \pland  \plneg p_4)$ \\
      %$(a \natfor b)$   & $\nateq$  & $(b \natrev a)$ \\	
      $\plneg\, (\plneg p_1 \pland \plneg p_2)$ & $\nateq$ & $(p_1 \plor p_2)$ \\ 
      \bottomrule
    \end{tabular}
    \caption{Examples of the type of statements used for training and testing. These are relations between
      well-formed formulae, computed in terms of sets of satisfying
      interpretation functions $\sem{\cdot}$.}\label{tab:plexs}
  \end{subtable}
  \caption{Natural logic relations over sentences of propositional logic.}  
  \label{prop-figure}
\end{table}

In our experiments, we randomly generate unique pairs 
of formulae and compute the relation that holds for each pair.
We discard pairs in which either statement is either a tautology or a
contradiction, for which the seven relations in
Table~\ref{b-table} are undefined. The resulting set of formula pairs is
then partitioned into 13 bins according the number of logical operators in
the longer of the two formulae. We then sample 20\% of each
bin for a held-out test set.

\begin{figure*}[t]
  \centering
  \includegraphics[width=5.75in]{decayfig.eps}
%  \includegraphics[width=5.3in]{recursion\string_pairwise.eps}
  \caption{Model performance on propositional logic, by expression size.}
%    The dashed line indicates that only expressions of size four or less appeared in the training data. \textbf{Bottom:} Semantically distinct formulae should have different
%    representations. As measured by Euclidean distance, only the TreeRNTN
%    achieves this for formulae containing more than a small number of
%    connectives (\ii{and} in this example). The TreeRNN quickly collapses the representations,
%    failing to capture the meaning contrasts.

  \label{prop-results}

\end{figure*}

If we do not implement any constraint that the two statements being
compared are similar in any way, then the generated data are dominated
by statements in which the two formulae refer to largely separate
subsets of the six variables, which means that the $\natind$ relation
is almost always correct.  In an effort to balance the distribution of
relation labels without departing from the basic task of modeling
propositional logic, we disallow individual pairs of statements from
referring to more than four of the six propositional variables.

In order to test the model's generalization to unseen structures, we discard
training examples with more than 4 connectives, yielding 60k short training examples,
and 21k test examples containing up to 12 connectives.
We evaluated both the TreeRNN and TreeRNTN on the resulting data, initializing the model parameters randomly,
including the vector representations of the six variables. 
As a baseline, we also train a summing neural network
which is largely identical to the TreeRNN, except that instead of learning a parametric composition function,
it simply sums the term vectors in each expression to compose them. Unlike the two models under study,
whose ability to accurately simulate the logic is an open question, this baseline does not use word order,
and is as such guaranteed to ignore some information that it would need to succeed perfectly.

% TODO: Insert baseline result.
\paragraph{Results} Fig.~\ref{prop-results} shows the relationship
between test accuracy and statement length. BASELINE BASELINE
 BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE BASELINE   We found that both 
models were able to perform well on unseen small test examples, 
with TreeRNN accuracy above
98\% and TreeRNTN accuracy above 99\% on formulae below length five.
Starting at size four, performance gradually falls with increasing
size. Training accuracy was 99.4\% for the TreeRNN and 99.8\% for the TreeRNTN.

The performance of these models on small unseen test examples
indicates that they learned correct approximations of the underlying
logic. These approximations are accurate enough to yield
correct answers when the composition layer is only applied a small
number of times, but that the error in the approximation grows with
increasing expression size, resulting in gradually dropping performance.

It appears that the TreeRNTN model overfit somewhat relative to the plain TreeRNN,
performing better on the training data and short examples, but worse on longer
examples. Two factors are likely involved in this: even with the lower dimension,
the TreeRNTN composition function has about eight times as many parameters as the
TreeRNN, and thus more capacity to find a solution involving a large number of narrow 
generalizations that work on the training data but don't generalize well. In addition,
the TreeRNTN used here has weaker L2 regularization than the 
TreeRNN ($\lambda = 0.0003$ vs. $0.001$). This
yielded better performance averaged across all bins, but may have facilitated this effect.

Though imperfect, model performance on the most difficult set of expressions is nonetheless striking. 
The weaker TreeRNTN classifies true examples of every class but $\nateq$ 
(which is rare in long examples, and occurs only once here) correctly the majority of the time.
% TODO: Better present examples.

% \mynote{In an effort to better understand why model performance decays, 
%we evaluated both models on pairs of long formulae involving
%binary connectives, to assess how well they distinguish representations for semantically distinct
%formulae. We found that only the TreeRNTN is able to learn
%substantially different representations for pairs of differing formulae like $(a \pland (a \pland a))$
% and $(a \pland (a \pland b))$ when the difference between the two is placed under multiple operators.
%The TreeRNN separates the bare symbols by a euclidean distance of 3.5, but this falls to less than one once the two are placed under three operators, and quickly approaches zero as depth increases.
% The TreeRNTN separates the bare symbols by 2.5, and this falls off much more gradually to 1.3 with twelve operators.}

% we discovered that this model looses information about in longer
% statements in a particularly problematic way. It appears that that
% model is unable to distinguish between two sentences when the only
% difference between those sentences is embedded within a very deep
% structure. We evaluated both models on sentences that differed in only
% one term, but for which that one term was embedded under a large
% number of conjunctions, such as the pair \ii{a (and a)} and \ii{a (and
%   b)}, or the pair \ii{a (and (a (and a)))} and \ii{a (and (a (and
%   b)))}. We then measured the Euclidean distance between the vector
% representations of the two sentences in each pair. Our findings are
% shown in Figure~\ref{prop-falloff}, and show that while the TreeRNTN can
% distinguish the two sentences well at every size that we test, the TreeRNN
% fails after a depth of approximately six.