### 2015-04-04 writing/S'15 misc/newmodel.tex
Relation handling cleanup.

\section{Jointly learning parsing and semantic composition}

% TODO: Graph figure

\begin{figure}[tp]
  \centering\small
  \begin{tikzpicture}
    \def\dx{18pt}
    \def\dy{30pt}
    \newcounter{i}
    \tikzstyle{highlight}=[fill=dandelion!15]
    \tikzstyle{comment}=[fill=blue!15]

    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,6*\dy) {$\vec{y}^{(1, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,5*\dy) {$\vec{y}^{(2, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,5*\dy) {$\vec{y}^{(2, 2)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-2*\dx,4*\dy) {$\vec{y}^{(3, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,4*\dy) {$\vec{y}^{(3, 2)}$};
    \stepcounter{i}\node[highlight]  (\arabic{i}) at (2*\dx,4*\dy) {$\vec{y}^{(3, 3)}$};
    \stepcounter{i}\node[highlight]  (\arabic{i}) at (-3*\dx,3*\dy) {$\vec{y}^{(4, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,3*\dy) {$\vec{y}^{(4, 2)}$};
    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,3*\dy) {$\vec{y}^{(4, 3)}$};
    \stepcounter{i}\node  (\arabic{i}) at (3*\dx,3*\dy) {$\vec{y}^{(4, 4)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-4*\dx,2*\dy) {$\vec{y}^{(5, 1)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-2*\dx,2*\dy) {$\vec{y}^{(5, 2)}$};
    \stepcounter{i}\node  (\arabic{i}) at (0*\dx,2*\dy) {$\vec{y}^{(5, 3)}$};
    \stepcounter{i}\node  (\arabic{i}) at (2*\dx,2*\dy) {$\vec{y}^{(5, 4)}$};
    \stepcounter{i}\node[highlight]  (\arabic{i}) at (4*\dx,2*\dy) {$\vec{y}^{(5, 5)}$};
    \stepcounter{i}\node  (\arabic{i}) at (-5*\dx,1*\dy) {My};
    \stepcounter{i}\node  (\arabic{i}) at (-3*\dx,1*\dy) {hovercraft};
    \stepcounter{i}\node  (\arabic{i}) at (-1*\dx,1*\dy) {is};
    \stepcounter{i}\node  (\arabic{i}) at (1*\dx,1*\dy) {full};
    \stepcounter{i}\node  (\arabic{i}) at (3*\dx,1*\dy) {of};
    \stepcounter{i}\node  (\arabic{i}) at (5*\dx,1*\dy) {eels};

    \stepcounter{i}\node[comment]  (\arabic{i}) at (4*\dx,4*\dy) {\textsc{compose}};
    \stepcounter{i}\node[comment]  (\arabic{i}) at (-5*\dx,3*\dy) {\textsc{copy-L}};
    \stepcounter{i}\node[comment]  (\arabic{i}) at (6*\dx,2*\dy) {\textsc{copy-R}};

    \tikzstyle{extra} = [->, draw=black, line width=.75pt]
    \tikzstyle{valid} = [->, draw=red, line width=1.25pt]

          \draw [valid] (2) -- (1);
          \draw [valid] (3) -- (1);

          \draw [valid] (4) -- (2);
          \draw [extra] (5) -- (2);
          \draw [valid] (5) -- (3);
          \draw [valid] (6) -- (3);

          \draw [valid] (7) -- (4);
          \draw [extra] (8) -- (4);
          \draw [valid] (8) -- (5);
          \draw [extra] (9) -- (5);
          \draw [valid] (9) -- (6);
          \draw [valid] (10) -- (6);

          \draw [valid] (11) -- (7);
          \draw [extra] (12) -- (7);
          \draw [valid] (12) -- (8);
          \draw [extra] (13) -- (8);
          \draw [valid] (13) -- (9);
          \draw [extra] (14) -- (9);
          \draw [valid] (14) -- (10);
          \draw [valid] (15) -- (10);

          \draw [valid] (16) -- (11);
          \draw [valid] (17) -- (11);
          \draw [extra] (17) -- (12);
          \draw [valid] (18) -- (12);
          \draw [extra] (18) -- (13);
          \draw [valid] (19) -- (13);
          \draw [extra] (19) -- (14);
          \draw [valid] (20) -- (14);
          \draw [extra] (20) -- (15);
          \draw [valid] (21) -- (15);
  \end{tikzpicture}

        \caption{The model structure used to compare \ii{turtle} and
          \ii{animal}.  Learned term representations are fed through
          either an NN or NTN comparison layer and then to a softmax
        classifier over the seven relations in Table~\ref{b-table}.}
  \label{sample-figure2}
\end{figure}

\begin{figure}[tp]
  \centering
    \footnotesize

	\newcommand{\labeledtreenode}[4][3]{\put(#2){\makebox(0,0){{\fcolorbox{black}{#4}{\makebox(#1,0.3){#3}}}}}}

	\newcommand{\textlabel}[4][3.5]{\put(#2){\makebox(0,0){{\makebox(#1,0.3){#3}}}}}

	\definecolor{lexcolor}{HTML}{F5F7C4}
	\definecolor{compositioncolor}{HTML}{BBEBFF}
	\definecolor{comparisoncolor}{HTML}{FFC895}
	\definecolor{softmaxcolor}{HTML}{A5FF8A}

	\setlength{\unitlength}{0.61cm}

	%\resizebox{3in}{!}{
	\begin{picture}(13.15, 5.5)

	  \labeledtreenode[2.4]{6.5,5}{$P(\sqsubset) = 0.8$}{softmaxcolor}  
	  \put(6.5,3.7){\vector(0,1){1}}  
	  \labeledtreenode[5]{6.5,3.4}{turtle \emph{vs.}~animal}{comparisoncolor}

	  \textlabel{2.75,5}{Softmax classifier}{black}
	  \textlabel{2.0,3.4}{\parbox{2cm}{Comparison N(T)N layer}}{black}

	  \textlabel{6.5,2.1}{Optional embedding}{black}
	  \textlabel{6.5,1.5}{transformation NN layers}{black}

	  \textlabel{6.5,0.5}{Learned term vectors}{black}

	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	  \put(1.5,0.75){\vector(0,1){0.85}}
	  \labeledtreenode{1.5,0.5}{turtle}{lexcolor}

	  \put(1.5,2.25){\vector(4,1){3.25}}
	  \labeledtreenode{1.5,1.9}{turtle}{compositioncolor}

	  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	  \put(11.5,0.75){\vector(0,1){0.85}}
	  \labeledtreenode{11.5,0.5}{animal}{lexcolor}

	  \put(11.5,2.25){\vector(-4,1){3.25}}
	  \labeledtreenode{11.5,1.9}{animal}{compositioncolor}

	\end{picture}
	%}
        \caption{The model structure used to compare \ii{turtle} and
          \ii{animal}.  Learned term representations are fed through
          either an NN or NTN comparison layer and then to a softmax
        classifier over the seven relations in Table~\ref{b-table}.}
  \label{sample-figure}
\end{figure}

To compute the vector representation for a sentence, we populate the bottom row of the structure $\vec{y}^{(N,1...N)}$ with the embedddings of each of the words. We then compute the representations of each row starting at the second row from the bottom ($N - 1$), and moving from left to right. Computing one feature vector requires two steps. In the first step, a classifier produces a distribution $\vec{o}$ over the operations \{\textsc{copy-left, copy-right, compose}\}, and in the second step, that distribution is used to compute a feature vector $\vec{y}$.

To compute the distribution $\vec{o{}}$

Choose an operation

row i
col j
amount of context C

To compute the vector representation for a sentence

\begin{equation} 
\label{TreeRNN}
\vec{o}^{(i,j)} = \text{softmax}(\mathbf{M_o} \colvec{7}
	{\vec{y}^{(i + 1, j - C + 1)}}
	{...}
	{\vec{y}^{(i + 1, j)}}
	{\vec{y}^{(i + 1, j + 1)}}
	{...}
	{\vec{y}^{(i + 1, j + C)}}
	{\vec{o}^{(i, j - 1)}}
	 + \vec{b}_o\,) \\
\end{equation}

\begin{equation}
\vec{y}^{(i,j)} = 
o^{(i, j)}_{1} \circ \vec{y^{(i + 1, j)}} +\\
o^{(i, j)}_{2} \circ \vec{y^{(i + 1, j + 1)}} +\\
o^{(i, j)}_{3} \circ \tanh(\mathbf{M_y} \colvec{2}{\vec{y}^{(i + 1, j)}}{\vec{y}^{(i + 1, j + 1)}} + \vec{b}_y)
\end{equation}

For true relation $\rho$, and true operations $\omega^{(i,j)}$ for one example

\begin{equation}
loss = \lambda\theta^2-\ln(r_\rho) - \frac{1}{(N - 1)^2} \sum_{i = 1}^{N - 1} \sum_{j = 1}^{N - 1} \ln(o_{\omega^{(i,j)}})
\end{equation}