### 2015-04-06 writing/S'15 misc/newmodel.tex
Optimizations, mostly around NL gradients

\section{Jointly learning parsing and semantic composition}

% TODO: Graph figure

\begin{figure}[tp]
  \centering\small
 	\input{lattice_fig.tex}

        \caption{The model structure used to compare \ii{turtle} and
          \ii{animal}.  Learned term representations are fed through
          either an NN or NTN comparison layer and then to a softmax
        classifier over the seven relations in Table~\ref{b-table}.}
  \label{sample-figure2}
\end{figure}

To compute the vector representation for a sentence, we populate the bottom row of the structure $\vec{y}^{(N,1...N)}$ with the embedddings of each of the words. We then compute the representations of each row starting at the second row from the bottom ($N - 1$), and moving from left to right. Computing one feature vector requires two steps. In the first step, a classifier produces a distribution $\vec{o}$ over the operations \{\textsc{copy-left, copy-right, compose}\}, and in the second step, that distribution is used to compute a feature vector $\vec{y}$.

To compute the distribution over operations at one node, $\vec{o}^{(i, j)}$,

Choose an operation

row i
col j
amount of context C

To compute the vector representation for a sentence

\begin{equation} 
\label{TreeRNN}
\vec{o}^{(i,j)} = \text{softmax}(\mathbf{M_o} \colvec{7}
	{\vec{y}^{(i + 1, j - C + 1)}}
	{...}
	{\vec{y}^{(i + 1, j)}}
	{\vec{y}^{(i + 1, j + 1)}}
	{...}
	{\vec{y}^{(i + 1, j + C)}}
	{\vec{o}^{(i, j - 1)}}
	 + \vec{b}_o\,) \\
\end{equation}

\begin{equation}
\vec{y}^{(i,j)} = 
o^{(i, j)}_{1} \circ \vec{y^{(i + 1, j)}} +\\
o^{(i, j)}_{2} \circ \vec{y^{(i + 1, j + 1)}} +\\
o^{(i, j)}_{3} \circ \tanh(\mathbf{M_y} \colvec{2}{\vec{y}^{(i + 1, j)}}{\vec{y}^{(i + 1, j + 1)}} + \vec{b}_y)
\end{equation}

For true relation $\rho$, and true operations $\omega^{(i,j)}$ for one example

\begin{equation}
loss = \lambda\theta^2-\ln(r_\rho) - \frac{1}{(N - 1)^2} \sum_{i = 1}^{N - 1} \sum_{j = 1}^{N - 1} \ln(o_{\omega^{(i,j)}})
\end{equation}