### 2015-05-26 writing/SNLI paper/model_structure_fig.tex
Lots of paper cleanup.

\begin{figure}[tp]
  \centering
\scalebox{0.85}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{29pt}

    \tikzstyle{label}=[text width=40mm,align=center]    
    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center]
    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center]
    \tikzstyle{e}=[fill=green!40,text width=25mm,align=center]
    \tikzstyle{m}=[draw=black,text width=38mm,align=center]

    \node[softmax]  (softmax) at (0*\dx,6*\dy) {3-way softmax classifier};
    \node[preclass]  (pc3) at (0*\dx,5*\dy) {100d $\tanh$ layer};
    \node[preclass]  (pc2) at (0*\dx,4*\dy) {100d $\tanh$ layer};
    \node[preclass]  (pc1) at (0*\dx,3*\dy) {100d $\tanh$ layer};
    \node[e]  (pe) at (-3*\dx,1.85*\dy) {50d premise};
    \node[e]  (he) at (3*\dx,1.85*\dy) {50d hypothesis};
    \node[m]  (pem) at (-3*\dx,0.5*\dy) {embedding model\\ with premise input};
    \node[m]  (hem) at (3*\dx,0.5*\dy) {embedding model\\ with hypothesis input};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

          \draw [fwd] (pc3) -- (softmax);
          \draw [fwd] (pc2) -- (pc3);
          \draw [fwd] (pc1) -- (pc2);
          \draw [fwd] (pe) -- (pc1);
          \draw [fwd] (he) -- (pc1);
          \draw [fwd] (hem) -- (he);
          \draw [fwd] (pem) -- (pe);

  \end{tikzpicture}}

        \caption{The classification architecture used in all NLI evaluations: for each model evaluated in Table~\ref{tab:nnresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 50d inputs shown here. In some transfer learning experiments (\S\ref{sec:transfer}), the premise embedding can be replaced with a zero vector for single-sentence tasks.}
  \label{modelstructure}
\end{figure}