### 2015-05-28 writing/SNLI paper/model_structure_fig.tex
Paper updates; config updates.

\begin{figure}[tp]
  \centering
\scalebox{0.85}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{29pt}

    \tikzstyle{label}=[text width=40mm,align=center]    
    \tikzstyle{softmax}=[fill=red!40,text width=40mm,align=center]
    \tikzstyle{preclass}=[fill=orange!40,text width=50mm,align=center]
    \tikzstyle{e}=[fill=green!40,text width=26mm,align=center]
    \tikzstyle{m}=[draw=black,text width=38mm,align=center]

    \node[softmax]  (softmax) at (0*\dx,6*\dy) {3-way softmax classifier};
    \node[preclass]  (pc3) at (0*\dx,5*\dy) {200d $\tanh$ layer};
    \node[preclass]  (pc2) at (0*\dx,4*\dy) {200d $\tanh$ layer};
    \node[preclass]  (pc1) at (0*\dx,3*\dy) {200d $\tanh$ layer};
    \node[e]  (pe) at (-3*\dx,1.85*\dy) {100d premise};
    \node[e]  (he) at (3*\dx,1.85*\dy) {100d hypothesis};
    \node[m]  (pem) at (-3*\dx,0.5*\dy) {embedding model\\ with premise input};
    \node[m]  (hem) at (3*\dx,0.5*\dy) {embedding model\\ with hypothesis input};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

          \draw [fwd] (pc3) -- (softmax);
          \draw [fwd] (pc2) -- (pc3);
          \draw [fwd] (pc1) -- (pc2);
          \draw [fwd] (pe) -- (pc1);
          \draw [fwd] (he) -- (pc1);
          \draw [fwd] (hem) -- (he);
          \draw [fwd] (pem) -- (pe);

  \end{tikzpicture}}

        \caption{The classification architecture used in all NLI evaluations: for each model evaluated in Table~\ref{tab:nnresults}, two identical copies of the model are run with the two sentences as input, and their outputs are used as the two 50d inputs shown here. In some transfer learning experiments (\S\ref{sec:transfer}), the premise embedding can be replaced with a zero vector for single-sentence tasks.}
  \label{modelstructure}
\end{figure}