### 2015-06-15 writing/S'15 LSTM/abstract.tex
Chris turn at abstract.

\begin{abstract}
%Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but 
%Tree-structured neural networks aim to deliver robust sentential meaning representations, but 
Tree-structured neural networks encode a particular tree geometry for a sentence in the network design.
However, these models have at best only slightly outperformed simpler sequence-based models.
% these models largely have not %outperformed simpler sequence-based models by substantial margins. 
% significantly outperformed simpler sequence-based models.
We hypothesize that neural sequence models like LSTMs are in fact able to discover and implicitly use %the same kinds of 
recursive compositional structure, %that the tree-structured ones are built around---
at least for tasks with clear cues to that structure in the data. 
% , mitigating the advantage of a tree-structured architecture. 
We demonstrate this possibility using an artificial data task for which recursive compositional structure is crucial, and find that the sequence model can learn the underlying patterning. The sequence model is better in that it learns the value of tree structure from the data in an emergent way, while the tree-structured model is better in being able to learn with greater statistical efficiency due to its informative prior model structure.
% structure, though it is less efficient at learning than the tree models, succeeding only after exposure to a larger and richer training set.
\end{abstract}